---
title: "Storing nested spatial output with tar_terra_nested(): a MESS example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Storing nested spatial output with tar_terra_nested(): a MESS example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Some functions return a *list* of `terra::SpatRaster` objects rather than a
single raster. The [`geotargets`](https://docs.ropensci.org/geotargets/)
package can serialize individual `SpatRaster` objects, but not a list that
contains several of them. `tar_terra_nested()` fills this gap by recursively
finding and serializing every spatial object inside an arbitrary R structure.

A natural example is `bssdm::mess()`, which calculates Multivariate
Environmental Similarity Surfaces. With `full = TRUE` it returns a
`MessResult` list of four `SpatRaster` layers: `mess`, `mess_by_variable`,
`mod`, and `mos`.

This vignette keeps things lightweight: rather than downloading global climate
data via `geodata`, it builds a small synthetic three-variable raster so the
example runs in a fraction of a second and requires no network access.

```{r}
library(targets)

tar_dir({

  # tar_dir() runs the pipeline in a temporary directory.

  tar_script({

    library(targets)
    library(geotargets)
    library(terra)
    library(bssdm)
    library(targets.utils)

    # Build a small synthetic climate raster and compute MESS against a
    # sample of reference cells. Everything is generated in-memory, so no
    # data is downloaded.
    make_mess <- function() {

      set.seed(4172)

      r <- rast(
        nrows = 20, ncols = 20, nlyrs = 3,
        xmin = 0, xmax = 10, ymin = 0, ymax = 10
      )
      names(r) <- c("bio1", "bio12", "bio5")
      values(r) <- cbind(
        runif(ncell(r), 5, 25),      # mean temperature
        runif(ncell(r), 200, 2000),  # annual precipitation
        runif(ncell(r), 20, 40)      # max temperature
      )

      # Reference values drawn from a random sample of cells
      ref <- spatSample(
        r, size = 30, method = "random",
        as.df = TRUE, na.rm = TRUE
      )

      # mess() returns a MessResult: a list of several SpatRasters
      mess(r, ref, full = TRUE)
    }

    list(
      # tar_terra_nested() handles the list of rasters that tar_target()
      # alone could not serialize.
      tar_terra_nested(
        mess_result,
        make_mess()
      )
    )
  })

  tar_make()

  # Read the target back: the MessResult list is reconstructed, with each
  # SpatRaster restored from its own file on disk.
  x <- tar_read(mess_result)
  print(class(x))
  print(names(x))
  x$mess
  
  library(bssdm)
  
  plot(x)

})
```

The reconstructed object retains its original `MessResult` class and all four
`SpatRaster` layers, each transparently restored from disk. The same approach
works for any function that returns spatial objects nested inside lists, data
frames, or other R structures.
